---
title: "Example of Central Limit Theorem using Exponential Distribution"
author: José Rodrigo Cervantes Polanco
output: pdf_document
---

## Overview

The **Central Limit Theorem** (CLT) is one of the most important theorems in statistics. For our purposes, the CLT states that the distribution of averages of I.I.D. random variables becomes that of a normal distribution as the sample size increases. In this project, we will use the exponential distribution as an example to understand the Central Limit Theorem.

## Central Limit Theorem (CLT)

Let $\{ X_{1} , \cdots , X_{n} \}$ be a sequence of I.I.D random variables of size $n$ drawn from distributions of expected values given by $\mu$ and finite variance given by $\sigma^2$. Suppose we are interested in the sample average
$$ \bar X_{n} = \frac{X_{1} + \cdots + X_{n}}{n} $$

of these random variables. The CLT states that as $n$ gets larger, the distribution of $\bar X_n$ is close to the normal distribution with mean $\mu$ and variance $\frac{\sigma^2}{n}$.


## Basic settings

We fixed a few basic settings to be used during this analysis. 
```{r echo=TRUE, cache=FALSE, message=FALSE}
setwd("~/Documents/ProgramsGitHub/StatInference")  # working directory
library(ggplot2)      # to be able to graph
set.seed(0)           # to be able to reproduce
```


## Simulations:

In this section we are going to do 1000 simulations of averages of 40 exponentials with $\lambda = 0.2$. 

```{r}
numsim = 1000                                                # Number of simulations
sample = NULL                                                # Initialize the simulations
for (i in 1 : numsim) sample = c(sample, mean(rexp(40,0.2))) # Create the simulations
sample<-as.data.frame(sample)                                # Create a dataframe (ggplot)
head(sample)                                                 # First simulations
```

## Sample Mean versus Theoretical Mean

Now, we are going to compare the sample mean versus the theoretical mean. We know that the mean of a random variable $X$ with exponential distribution, is given by $\mu(X)=1/\lambda$. In our simulations $\lambda=0.2$, therefore $\mu(X)=5$. A consequence of the CLT is that the mean of the random variable given by the average of $40$ exponentials is approximated by the mean of the exponential distribution, in this project $5$.

Fisrt, we graph both the sample mean and the theoretical mean:
```{r, tidy=FALSE}
g <- ggplot(sample,aes(sample))     # Enter the database to be graph
g <- g + geom_histogram(binwidth=.1,aes( y=..density..))  # Type of graph / histogram
g <- g + geom_vline(aes(xintercept=5, color="Theoretical Mean"), 
                    linetype="solid", size=0.5, show_guide = TRUE)
g <- g + geom_vline(aes(xintercept=mean(sample), color="Sample Mean"), 
                    linetype="solid", size=0.5,show_guide = TRUE) 
g <- g + scale_colour_manual("", values = c("Sample Mean" = "green", 
                                            "Theoretical Mean" ="red" ))
g <- g + labs(title = "Sample Mean vs Theoretical Mean",x="Sample",y="Frequency")
g
```

Second, we calculate the sample mean:
```{r}
mean(sample$sample)
```
From this calculation and the previous graph, we have checked that $5$ is a good approximation for the sample mean. 

## Sample Variance versus Theoretical Variance:

In this part, we are going to compare the sample variance versus the theoretical variance. We know that the variance of a random variable $X$ with exponential distribution, is given by $\text{var}(X)=1/\lambda^2$. In our simulations $\lambda=0.2$, therefore $\text{var}(X)=25$. A consequence of the CLT is that the variance of the random variable given by the average of $40$ exponentials is approximated by the variance of the exponential distribution divide by $40$, in this project $25/40 = 0.625$.

Fisrt, we graph bands with a width of $1$ sample standard deviation and $1$ theoretical standard deviation (remember that the standard deviation is the square root of the variance):
```{r, tidy=FALSE}
g <- ggplot(sample,aes(sample))     # Enter the database to be graph
g <- g + geom_histogram(binwidth=.1,aes( y=..density..))  # Type of graph / histogram
g <- g + geom_vline(aes(xintercept=5+(25/40)^(1/2), color="1 - Theoretical SD"), 
                    linetype="solid", size=0.5, show_guide = TRUE)
g <- g + geom_vline(aes(xintercept=5-(25/40)^(1/2), color="1 - Theoretical SD"), 
                    linetype="solid", size=0.5, show_guide = TRUE)
g <- g + geom_vline(aes(xintercept=mean(sample)+sd(sample), color="1 - Sample SD"), 
                    linetype="solid", size=0.5,show_guide = TRUE) 
g <- g + geom_vline(aes(xintercept=mean(sample)-sd(sample), color="1 - Sample SD"), 
                    linetype="solid", size=0.5,show_guide = TRUE)
g <- g + scale_colour_manual("", values = c("1 - Sample SD" = "green", 
                                            "1 - Theoretical SD" ="red" ))
g <- g + labs(title = "Sample SD vs Theoretical SD",x="Sample",y="Frequency")
g
```

Second, we calculate the sample mean:
```{r}
var(sample$sample)
```
From this calculation and the previous graph, we have checked that $25/40 = 0.625$ is a good approximation for the sample variance. 


## Distribution:

The way to think about the CLT is that $\bar X_n$ is approximately $N(\mu, \sigma^2 / n)$, where $\bar X_n$ is the average of I.I.D. random variables, $n$ the number of random variables,  $\mu$ is the mean and $\sigma$ the variance of one of those random variables. In our  

Finally, we graph a normal distribution with mean equals $4.99$ and variance equals $0.6182$ over the  histogram of the frequencies of the sample that we have been using in this project: 

```{r, tidy=FALSE}
g <- ggplot(sample,aes(sample))     # Enter the database to be graph
g <- g + geom_histogram(binwidth=.1,aes( y=..density..))  # Type of graph / histogram
g <- g + stat_function(fun = dnorm,args = list(mean=mean(sample$sample), 
                                               sd = sd(sample$sample)), color = "green")
g <- g + labs(title="Normal Distribution Approximation",x="Sample",y="Frequency")
g
```

The last graph shows that the distribution of our sample of $1000$ averages of $40$ exponentials with $\lambda=0.2$ can be approximated by a normal distribution with mean $4.99$ (approx $5$) and variance $0.6182$ (approx $0.625$).  